\documentclass[12pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
\author{Henry Yu}
\begin{document}

\raggedright

\begin{center}
	\huge{Note 1}
\end{center}

\section*{Propositional Logic}

To become fluent in working with mathematical statements, you need to understand the basic framework of the language of mathematics.

\bigbreak

\textbf{Proposition}: a statement which is either true or false

\bigbreak

These statements are propositions:
\begin{itemize}
	\item[1.] $\sqrt{3}$ is irrational.
	\item[2.]$1+1=5$.
\end{itemize}

These statements are not propositions:
\begin{itemize}
	\item[1.] $2+2$.
	\item[2.]$x^2+3x=5$. (since x is unknown)
\end{itemize}

Propositions should not include fuzzy terms, so these statements aren't propositions either (although some sources may say they are):
\begin{itemize}
	\item[1.] Arnold Schwarzenegger often eats broccoli. ("often" is fuzzy)
	\item[2.] Henry VIII was unpopular. ("unpopular" is fuzzy)
\end{itemize}

Propositions may be joined together to form more complex statements. Let $P$, $Q$, \text{and } $R$ be variables representing propositions. The simplest way of joining these propositions together is to use the connectives "and," "or," and "not."
\begin{itemize}
	\item[1.] \textbf{Conjunction}: $P\wedge Q$ ("$P$ and $Q$"). True only when both $P$ and $Q$ are true.
	\item[2.] \textbf{Disjunction}: $P\vee Q$ ("$P$ or $Q$"). True when at least one of $P$ and $Q$ is true.
	\item[3.] \textbf{Negation}: $\neg P$ ("not $P$"). True when $P$ is false.
\end{itemize}

Statements with variables (like these) are called \textbf{propositional forms}.

\bigbreak

A fundamental principle known as the \textbf{law of the excluded middle} says that, for any proposition $P$, either $P$ is true or $\neg P$ is true (but not both). Thus $P\vee\neg P$ is always true, regardless of the truth value of $P$. A propositional form is always true regardless of its truth values is called a \textbf{tautology}. Conversely, a statement such as $P\wedge\neg P$, which is always false, is called a \textbf{contradiction}.

\bigbreak

A \textbf{truth table} is used to describe the possible values of a propositional form. Truth tables are the same as function tables: you list all possible input values for the variables, and then list the outputs given for those inputs.

\bigbreak

Here are the truth tables for conjunction, disjunction, and negation:
\begin{center}
	\begin{tabular}{ |c|c| }
		\hline
		$P$ & $Q$ \\
		\hline
		T & T \\
		\hline
		T & F \\ 
		\hline
		F & T \\ 
		\hline
		F & F \\
		\hline
	\end{tabular}
	\begin{tabular}{ |c| }
		\hline
		$P\wedge Q$ \\
		\hline
		T \\
		\hline
		F \\
		\hline
		F \\
		\hline
		F \\
		\hline
	\end{tabular}
	\begin{tabular}{ |c| }
		\hline
		$P\vee Q$ \\
		\hline
		T \\
		\hline
		T \\
		\hline
		T \\
		\hline
		F \\
		\hline
	\end{tabular}
	\begin{tabular}{ |c|c| }
		\hline
		$P$ & $\neg P$ \\
		\hline
		T & F \\
		\hline
		F & T \\
		\hline		
	\end{tabular}
\end{center}

\newpage

The most important and subtle propositional form is an \textbf{implication}:
\begin{itemize}
	\item[4.] \textbf{Implication}: $P\Rightarrow Q$ ("$P$ implies $Q$"). This is the same as "if $P$, then $Q$."
\end{itemize}

Here, $P$ is called the \textbf{hypothesis} of the implication, and $Q$ is the \textbf{conclusion}.

\bigbreak

Here are some examples:
\begin{itemize}
	\item If you stand in the rain, then you'll get wet.
	\item If you passed the class, you received a certificate.
\end{itemize}

An implication $P\Rightarrow Q$ is false only when $P$ is true and $Q$ is false.

\bigbreak

Here is the truth table for $P\Rightarrow Q$ (along with an additional column):
\begin{center}
	\begin{tabular}{ |c|c|c|c| }
		\hline
		$P$ & $Q$ & $P\Rightarrow Q$ & $\neg P \vee Q$ \\
		\hline
		T & T & T & T \\
		\hline
		T & F & F & F \\
		\hline
		F & T & T & T \\
		\hline
		F & F & T & T \\ 
		\hline
	\end{tabular}
\end{center}

Note the $P\Rightarrow Q$ is always true when $P$ is false. When an implication is stupidly true because the hypothesis is false, we say that it is \textbf{vacuously true}.

\bigbreak

Note also that $P\Rightarrow Q$ is \textbf{logically equivalent} to $\neg P \vee Q$. We write this as $(P\Rightarrow Q)\equiv(\neg P \vee Q)$.

\bigbreak

Here are some different ways of meaning $P\Rightarrow Q$:
\begin{itemize}
	\item[1.] if $P$, then $Q$
	\item[2.] $Q$ if $P$
	\item[3.] $P$ only if $Q$
	\item[4.] $P$ is sufficient for $Q$
	\item[5.] $Q$ is necessary for $P$
	\item[6.] $Q$ unless not $P$
\end{itemize}

If both $P\Rightarrow Q$ and $Q\Rightarrow P$ are true, then we say "$P$ if and only if $Q$" (abbreviated "$P$ iff $Q$"). Formally, we write $P\Leftrightarrow Q$. Note that $P\Leftrightarrow Q$ is true when $P$ and $Q$ have the same truth values (both true or false).

\bigbreak

Given an implication $P\Rightarrow Q$, we can also define its
\begin{itemize}
	\item[1.] \textbf{Contrapositive}: $\neg Q\Rightarrow \neg P$
	\item[2.] \textbf{Converse}: $Q\Rightarrow P$
\end{itemize}

\newpage

Here are some truth tables:

\begin{center}
	\begin{tabular}{ |c|c|c|c||c|c|c|c| }
		\hline
		$P$ & $Q$ & $\neg P$ & $\neg Q$ & $P\Rightarrow Q$ & $Q\Rightarrow P$ & $\neg Q\Rightarrow \neg P$ & $P\Leftrightarrow Q$ \\
		\hline
		T & T & F & F & T & T & T & T \\
		\hline
		T & F & F & T & F & T & F & F \\
		\hline
		F & T & T & F & T & F & T & F \\
		\hline
		F & F & T & T & T & T & T & T \\
		\hline
	\end{tabular}
\end{center}

Note that $P\Rightarrow Q$ and its contrapositive have the same truth values, so they are logically equivalent: $(P\Rightarrow Q)\equiv(\neg Q\Rightarrow\neg P)$.

\bigbreak

Also note that $P\Rightarrow Q$ and $Q\Rightarrow P$ are not logically equivalent.

\section*{Quantifiers}

The mathematical statements you'll see in practice will look something like this:
\begin{itemize}
	\item[1.] For all natural numbers $n$, $n^2+n+41$ is prime.
	\item[2.] If $n$ is an odd integer, so is $n^3$.
	\item[3.] There is an integer $k$ that is both even and odd.
\end{itemize}

These statements assert something about lots of simple propositions all at once. For instance, the first statement is asserting that $0^2+0+41$ is prime, $1^2+1+41$ is prime, and so on.

\bigbreak

The last statement says that as $k$ ranges over all possible integers, we will find at least one value of $k$ for which the statement is satisfied.

\bigbreak

Compared to the previous statement "$x^2+3x=5$," these examples are quantified over a "universe." To express these statements mathematically, we need two \textbf{quantifiers}:
\begin{itemize}
	\item[1.] the universal quantifier $\forall$ ("for all")
	\item[2.] the existential quantifier $\exists$ ("there exists")
\end{itemize}

Examples:
\begin{itemize}
	\item[1.] "Some mammals lay eggs."
\end{itemize}

Mathematically, "some" means "at least one," so the statement is actually saying "There exists a mammal $x$ such that $x$ lays eggs." If we let our universe $U$ to be the set of mammals, then we can write: $(\exists x\in U)(x \text{ lays eggs})$.

\begin{itemize}
	\item[2.] "For all natural numbers $n$, $n^2+n+41$ is prime."
\end{itemize}

In this case, our universe becomes the set of natural numbers, denoted as $\mathbb{N}$: $(\forall n \in \mathbb{N})(n^2+n+41\text{ is prime})$.

\bigbreak

We refer to a statement which refers to a variable as a \textbf{predicate} or as a \textbf{propositional formula} when replacing the variable with a value makes the statement true or false.

\newpage

Note that in a finite universe, we can express existentially and universally quantified propositions without quantifiers, using disjunctions and conjunctions respectively.

\bigbreak

For example, if our universe $U$ is $\{1,2,3,4\}$, then \\
$(\exists x \in U)P(x)$ is logically equivalent to $P(1)\vee P(2)\vee P(3)\vee P(4)$, and \\
$(\forall x \in U)P(x)$ is logically equivalent to $P(1)\wedge P(2) \wedge P(3) \wedge P(4)$.
 
\bigbreak
 
However, in an infinite universe, this is not possible.

\section*{Negation}

Let's look at how to negate conjunctions and disjunctions:
\begin{center}
 	$\neg(P\wedge Q)\equiv(\neg P\vee\neg Q)$ \\
 	$\neg(P\vee Q)\equiv(\neg P\wedge\neg Q)$
\end{center}

These two equivalences are known as \textbf{De Morgan's Laws}, and they are quite intuitive: for example, if it is not the case that $P\wedge Q$ is true, then either $P$ or $Q$ is false (and vice versa).

\bigbreak

Negating propositions involving quantifiers actually follows analogous laws.

\bigbreak
Assume that the universe is $\{1,2,3,4\}$ and let $P(x)$ denote the propositional formula "$x^2>10$". \\
Check that $\exists xP(x)$ is true but $\forall xP(x)$ is false. \\
Observe that both $\neg(\forall xP(x))$ and $\exists x\neg P(x)$ are true because $P(1)$ is false. \\
Also note that both $\forall x\neg P(x)$ and $\neg(\exists xP(x))$ are false, because $P(4)$ is true.
\end{document}