\documentclass[12pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
\begin{document}

\raggedright

\section*{Questionnaire 2}

\begin{enumerate}
\item Provide an example of where the bear classification model might work poorly in production, due to structural or style differences in the training data. \\
\item Where do text models currently have a major deficiency? \\
\item What are possible negative societal implications of text generation models? \\
\item In situations where a model might make mistakes, and those mistakes could be harmful, what is a good alternative to automating a process? \\
\item What kind of tabular data is deep learning particularly good at? \\
\item What's a key downside of directly using a deep learning model for recommendation systems? \\
\item What are the steps of the Drivetrain Approach? \\
\item How do the steps of the Drivetrain Approach map to a recommendation system? \\
\item Create an image recognition model using data you curate, and deploy it on the web. \\
\item What is DataLoaders? \\
\item What four things do we need to tell fastai to create DataLoaders? \\
\item What does the splitter parameter to DataBlock do? \\
\item How do we ensure a random split always gives the same validation set? \\
\item What letters are often used to signify the independent and dependent variables? \\
\item What's the difference between the crop, pad, and squish resize approaches? When might you choose one over the others? \\
\item What is data augmentation? Why is it needed? \\
\item What is the difference between item\_tfms and batch\_tfms? \\
\item What is a confusion matrix? \\
\item What does export save? \\
\item What is it called when we use a model for getting predictions, instead of training? \\
\item What are IPython widgets? \\
\item When might you want to use CPU for deployment? When might GPU be better? \\
\item What are the downsides of deploying your app to a server, instead of to a client (or edge) device such as a phone or PC? \\
\item What are three examples of problems that could occur when rolling out a bear warning system in practice? \\
\item What is "out-of-domain data"? \\
\item What is "domain shift"? \\
\item What are the three steps in the deployment process? \\
\end{enumerate}

\section*{Further Research}

\begin{enumerate}
\item Consider how the Drivetrain Approach maps to a project or problem you're interested in. \\
\item When might it be best to avoid certain types of data augmentation? \\
\item For a project you're interested in applying deep learning to, consider the thought experiment "What would happen if it went really, really well?" \\
\item Start a blog, and write your first blog post. For instance, write about what you think deep learning might be useful for in a domain you're interested in. \\
\end{enumerate}

\end{document}